---
title: "Learning machine final assignment"
author: "Juan Villarreal"
date: "November 11, 2017"
output: html_document
---


```{r setup, include=FALSE, echo=TRUE}
knitr::opts_chunk$set(echo = TRUE)

```

## Loading data and libraries.


```{r loading, echo=TRUE}
download.file('https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv', 'pml-training.csv')
download.file('https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv', 'pml-testing.csv')
pmltest <- read.csv("pml-testing.csv")
pmltrain <- read.csv("pml-training.csv")
library(ggplot2)
library(caret)
library(dplyr)
library(parallel)
library(doParallel)
set.seed(3261)
```


# Executive summary

This document contains the analysis on Weight Lifting Exercises Dataset, Groupware have been very generous allowing their data to be used for this assignment.

Training data set belongs to Six young health participants that performed one set of 10 repetitions of the Unilateral Dumbbell Biceps Curl in five different fashions: 

## Exactly according to the specification (Class A).
## Throwing the elbows to the front (Class B).
## Lifting the dumbbell only halfway (Class C).
## Lowering the dumbbell only halfway (Class D).
## Throwing the hips to the front (Class E).

Read more: http://groupware.les.inf.puc-rio.br/har#ixzz4y9VpkiRS


The intention of this analysis is to predict the Class (A,B,C,D,E) using learning machine algorithm with provided training and testing dataset from Human Activity Recognition.


# Exploratory Analysis on training and testing data sets.


```{r exploratory, echo=TRUE}
arrange(distinct(pmltrain,user_name,cvtd_timestamp,classe), user_name, cvtd_timestamp,classe)
```

Observing training dataset in detail, the classification by each participant and the timestamp when they made the excersise, my conclusion with this quick review is, timestamp is not associated with the classification, "classe"" is the column that tells "how well" they performed the excersice, we don't need to know "when", this will impact on accuracy while applying the predict function heance is going to be remove the 3 columns associated with timestamp(raw_timestamp_part_1, raw_timestamp_part_2, cvtd_timestamp), also the X column that is just a sequence number and is not importan for the regression model. 

```{r exploratory2, echo=TRUE}
arrange(distinct(pmltest, user_name, cvtd_timestamp, new_window), user_name, cvtd_timestamp, new_window)
```

Observing test dataset in detail, we can see that testing data set is trying to predict one or more times the same individual, with different results, I will discard column X which is a sequence number, raw_timestamp_part_1, raw_timestamp_part_2, cvtd_timestamp to be consistent with training data set.


Other relevant observation is the the testing data set is going to predict base on new_windows = "no", that means we can ignore or delete the values for new_windows = "yes" in the training data set, also filtering the values will allow to delete several columns witn useless statistical values like a "tab" character, blank spaces, or NA values, this will increase the performance of the model and training and testing data sets will not be affected, the valiables with the relevant invormation for this excersice will keep intact.



```{r tidyng_data, echo=TRUE}


pmltrain_tidy <- subset(pmltrain, new_window == "no")
pmltrain_tidy [,c("X","raw_timestamp_part_1","raw_timestamp_part_2","cvtd_timestamp","new_window", "num_window")] <- list(NULL)
pmltrain_tidy [,c("kurtosis_roll_belt","kurtosis_picth_belt", "kurtosis_yaw_belt","skewness_roll_belt","skewness_roll_belt.1" ,
"skewness_yaw_belt","max_yaw_belt", "min_yaw_belt", "amplitude_yaw_belt" ,	"kurtosis_roll_arm" , "kurtosis_picth_arm","kurtosis_yaw_arm","skewness_roll_arm" ,"skewness_pitch_arm"  , "skewness_yaw_arm" ,
"kurtosis_roll_dumbbell","kurtosis_picth_dumbbell","kurtosis_yaw_dumbbell","skewness_roll_dumbbell","skewness_pitch_dumbbell",
"skewness_yaw_dumbbell","max_yaw_dumbbell", "min_yaw_dumbbell","amplitude_yaw_dumbbell","kurtosis_roll_forearm","kurtosis_picth_forearm","kurtosis_yaw_forearm",
 "skewness_roll_forearm" ,"skewness_pitch_forearm","skewness_yaw_forearm", "max_yaw_forearm", "min_yaw_forearm",
"amplitude_yaw_forearm")] <- list(NULL)
pmltrain_tidy[sapply(pmltrain_tidy, function(x) all(is.na(x)))] <- NULL



pmltest_tidy <- subset(pmltest, new_window == "no")
pmltest_tidy [,c("X","raw_timestamp_part_1","raw_timestamp_part_2","cvtd_timestamp","new_window", "num_window")] <- list(NULL)
pmltest_tidy [,c("kurtosis_roll_belt","kurtosis_picth_belt", "kurtosis_yaw_belt","skewness_roll_belt","skewness_roll_belt.1" ,
"skewness_yaw_belt","max_yaw_belt", "min_yaw_belt", "amplitude_yaw_belt" ,	"kurtosis_roll_arm" , "kurtosis_picth_arm","kurtosis_yaw_arm","skewness_roll_arm" ,"skewness_pitch_arm"  , "skewness_yaw_arm" ,
"kurtosis_roll_dumbbell","kurtosis_picth_dumbbell","kurtosis_yaw_dumbbell","skewness_roll_dumbbell","skewness_pitch_dumbbell",
"skewness_yaw_dumbbell","max_yaw_dumbbell", "min_yaw_dumbbell","amplitude_yaw_dumbbell","kurtosis_roll_forearm","kurtosis_picth_forearm","kurtosis_yaw_forearm",
 "skewness_roll_forearm" ,"skewness_pitch_forearm","skewness_yaw_forearm", "max_yaw_forearm", "min_yaw_forearm",
"amplitude_yaw_forearm")] <- list(NULL)
pmltest_tidy[sapply(pmltest_tidy, function(x) all(is.na(x)))] <- NULL
names(pmltrain_tidy)
names(pmltest_tidy)

```


We can see we have same columns in both data sets except by problem_id in the test data set, we will exclude it for the model.



```{r model, echo=TRUE}
cluster <- makeCluster(detectCores() - 1) # detecting how many cores the system has.
registerDoParallel(cluster)   # taking advantage on doParallel pakcage.
fitControl <- trainControl(method = "cv",number = 10, allowParallel = TRUE) 
modFit <- train(classe ~ . , method = "rf", data=pmltrain_tidy, trControl = fitControl) #fitting the model takes around 1 hour with i7 8GB RAM
modFit
```

Random Forest method is displaying 99.4% accuracy, we will predict with that method.



```{r prediction, echo=TRUE}
pred <- predict(modFit, subset(pmltest_tidy, select = -problem_id))
pmltest_tidy$predictionClasse <- pred
subset(pmltest_tidy, select = c(problem_id, user_name, predictionClasse))
```


## Conclusion
Radom Forest method is good enought to predict the classe of the fitbit information.


## Appendix



